Multisymplectic observables and higher Courant algebroids

### 2024-02-26 msgauge-Caprarola.tex
Slides Caprarola - slides done - narrative todo.

%- D0cum3nt ----------------------------------------------------------------------------------------------
\documentclass[10pt,handout]{beamer}

%- Packages -------------------------------------------------------------------------------------------
\usepackage{custom-style}
\usepackage{multicol}
\usepackage{stmaryrd}
\usepackage{animate}
\usetikzlibrary{positioning, arrows}
\usetikzlibrary{shapes}
\usepackage{fontawesome}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Why not a direct vertical arrow?}\label{frame:whyNoVertArrow}
	\begin{center}
			\includestandalone[width=.7\textwidth]{Pictures/Frame_BigDiagram_verticalArrow}
	\end{center}
	\vfill
	\begin{itemize}
		\item<2-> Assume exists \alert<2>{$\chi:C^\infty(M)_\omega \to C^\infty(M)_{\widetilde{\omega}}$} ~s.t. the square commutes.
				\begin{displaymath}
					\begin{cases}
						\chi(f) =& f - \iota_{\vHam_f} B
						\\
						\widetilde{\vHam}_{\chi(f)} =& \vHam_f
					\end{cases}
				\end{displaymath}
		\item<3-> Then
			\begin{align*}
				d \chi(f) &= - \iota_{\vHam_f}\omega - d \iota_{\vHam_f} B \\
				\parallel & \\
				-\iota_{\widetilde{\vHam}_{\chi(f)}} \widetilde{\omega} &= -\iota_{\vHam_f} \omega + \iota_{\vHam_f} d B
			\end{align*}
		\item<4-> I.e. $$\mathcal{L}_{\vHam_f} B = 0 ~ \forall f \in C^\infty(M) \qquad \qquad \text{\alert{contradiction!}}$$
	\end{itemize}
	\hfill \hyperlink{frame:pentagonMotivation}{\color{gray}$[<<]$}%\beamerskipbutton{}}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Pentagonal diagram in the \emph{prequantizable} case}\label{frame:pentagonPrequantum}
	Assume $\omega$ is \alert{prequantizable}:
	\pause
	%
		\begin{columns}
			\begin{column}{.6\linewidth}
				\begin{itemize}
					\item<2-> $\pi:P\to M$ \small associated $S^1$ bundle;
					\item<3-> $E\in \mathfrak{X}(P)$ \small fundamental v.f. of $S^1\curvearrowright P$;
					\item<4-> $\theta\in \Omega^1(P)^{S^1}$ \small connection $1$-form  s.t. $d \theta = \pi^\ast \omega$.
				\end{itemize}
			\end{column}
			\begin{column}{.4\linewidth}
				\begin{center}
					\includestandalone[width=1\textwidth]{Pictures/Frame_prequantum}
				\end{center}	
			\end{column}
		\end{columns}
	\pause \pause \vfill
	The Lie algebra map
	\begin{displaymath}
		\big(C^\infty(M),\{\cdot,\cdot\}_\omega\big) \xrightarrow{\Psi_\omega} 
		\big(\Gamma(TM\oplus\mathbb{R}),[\cdot,\cdot]_\omega\big)
	\end{displaymath}
	%
	\vfill
	factors through the Lie algebra of \emph{\alert{infinitesimal quanto-morphisms}}
	%
	\pause
	\begin{displaymath}
		Q(P,\theta)= \left \lbrace 
			Y \in \mathfrak{X}(P) 
			~ \Big\vert ~
			\mathcal{L}_Y \theta = 0
			\right \rbrace~.
	\end{displaymath}

@@

@@

	Consider a \emph{\bf{first-order Lagrangian Field Theory}}:
	\begin{itemize}
		\item<1-> {\color{blue} Kinematics} is encoded by a {\color{blue} \emph{configuration bundle}};
		\item<2-> {\color{red} Dynamics} is encoded by a {\color{red} \emph{Lagrangian density}};
		\item<3-> The {\color{orange} multicontangent bundle} is naturally $n$-plectic with $\omega := d \theta$ and \\
		$\theta \in \Omega^n(\Lambda)$ the tautological $n$-form
		{\small
		\begin{displaymath}
			\left[ \iota_{u_1 \wedge \ldots \wedge u_n} \theta \right]_\eta 
			= \iota_{u_1 \wedge \ldots \wedge u_n} \pi^\ast \eta 
			\qquad \qquad \forall \eta \in \wedge^n T^* E \, , \: \forall u_i \in T_\eta (\wedge^n T^* E) 		
		\end{displaymath}
		}
		\item<4-> The multisymplectic $(n+1)$-form is pulled back to $J^1 E$ through the \emph{Legendre transformation} $\mathbb{F}\Lag: J^1 E \to (J^1 E)^\ostar$
    given intrinsically by
    {\small
		\begin{displaymath}
        		\Big\langle \mathbb{F}\mathcal{L}(\gamma),\gamma'\Big\rangle=
        		\mathcal{L}\Big(\gamma\Big) + \frac{ d}{ d \epsilon}\mathcal{L}\Big(\gamma +\epsilon(\gamma'-\gamma)\Big)\Bigg\vert_{\epsilon=0} \qquad \forall \gamma,\gamma' \in J^1 E
    		\end{displaymath}
    }

    %
	\end{itemize}
	%
	\vfill
	\hfill \hyperlink{frame:hystMS}{\color{gray}$[<<]$}%\beamerskipbutton{}}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[t]{Vinogradov Algebroids}
	\begin{defblock}[Vinogradov algebroid (higher Courant)]
		\includestandalone[width=0.95\textwidth]{Pictures/Frame_vinogradov}		
	\end{defblock}
	\vfill
	\begin{itemize}
		\item<6-> $(n=1) ~ \Rightarrow$ standard twisted \emph{Lie algebroid};
		\item<7-> $(n=2) ~ \Rightarrow$ standard twisted \emph{Courant algebroid};
	\end{itemize}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Vinogradov $L_\infty$-algebra}
	Vin. alg.oids 
	$\quad\simeq\quad$ 	
	$NQ$-manifolds {\small($L_\infty$-algebroids)}
	$\quad\Rightarrow\quad$ 
	{\small Associated} $L_\infty$-algebra
	%
	\pause
	\begin{defblock}[Vinogradov $L_\infty$-algebra (Zambon) \cite{Zambon2012}]
		\includestandalone[width=0.95\textwidth]{Pictures/Frame_vinogradov-Linfty}	
	\end{defblock}
	\vfill
	\only<6->{
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=gray!1,draw=gray!30,anchor=base]            
            	 (base) at ($(current page.south)+(0,.5)$) [rotate=-0,text width=10cm,align=center] { \footnotesize{\color{gray}{
            	 $e_i = \binom{X_i}{\alpha_i} \in \mathfrak{X}(M)\oplus \Omega^{n-1}(M)$ 
            	 \quad~,\qquad
            	 $f_i \in \bigoplus_{k=0}^{n-2}\Omega^k(M)$.
            	 }}};
		}			
	}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{TODO}
	Details sulla ricetta di Zambon per costruire questa Lie infinito algebra.
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

	}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{TODO}
	OUTLOOK. La big picture. 
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

	\begin{frame}{}
	\label{frame:thankyouslide}
		\vfill
	  \centering 
	  {\Huge\color{red} 
	  \emph{Thank you for your attention!}}
		\vfill
		%
		\centering
		\fbox{\includegraphics[width=.4\textwidth]{Pictures/arxiv-cover}}
	\end{frame}

%------------------------------------------------------------------------------------------------
% APPENDIX
%------------------------------------------------------------------------------------------------
\appendix

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Appendix}
%------------------------------------------------------------------------------------------------

>>>>

%- D0cum3nt ----------------------------------------------------------------------------------------------
\documentclass[10pt]{beamer}

%- Packages -------------------------------------------------------------------------------------------
\usepackage{custom-style}
\usepackage{multicol}
\usepackage{stmaryrd}
\usepackage{animate}
\usetikzlibrary{positioning, arrows}
\usetikzlibrary{shapes}
\usepackage{fontawesome}
\usepackage{amsfonts}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Why not a direct vertical arrow?}\label{frame:whyNoVertArrow}
	\begin{center}
			\includestandalone[width=.7\textwidth]{Pictures/Frame_BigDiagram_verticalArrow}
	\end{center}
	\vfill
	\begin{itemize}
		\item<2-> Assume exists \alert<2>{$\chi:C^\infty(M)_\omega \to C^\infty(M)_{\widetilde{\omega}}$} ~s.t. the square commutes.
				\begin{displaymath}
					\begin{cases}
						\chi(f) =& f - \iota_{\vHam_f} B
						\\
						\widetilde{\vHam}_{\chi(f)} =& \vHam_f
					\end{cases}
				\end{displaymath}
		\item<3-> Then
			\begin{align*}
				d \chi(f) &= - \iota_{\vHam_f}\omega - d \iota_{\vHam_f} B \\
				\parallel & \\
				-\iota_{\widetilde{\vHam}_{\chi(f)}} \widetilde{\omega} &= -\iota_{\vHam_f} \omega + \iota_{\vHam_f} d B
			\end{align*}
		\item<4-> I.e. $$\mathcal{L}_{\vHam_f} B = 0 ~ \forall f \in C^\infty(M) \qquad \qquad \text{\alert{contradiction!}}$$
	\end{itemize}
	\hfill \onslide<4->{\hyperlink{frame:pentagonMotivation}{\color{gray}$[<<]$}}%\beamerskipbutton{}}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Pentagonal diagram in the \emph{prequantizable} case}\label{frame:pentagonPrequantum}
	Assume $\omega$ is \alert{prequantizable}:
	\pause
	%
		\begin{columns}
			\begin{column}{.6\linewidth}
				\begin{itemize}
					\item<2-> $\pi:P\to M$ \small associated $S^1$ bundle;
					\item<3-> $E\in \mathfrak{X}(P)$ \small fundamental v.f. of $S^1\curvearrowright P$;
					\item<4-> $\theta\in \Omega^1(P)^{S^1}$ \small connection $1$-form  s.t. $d \theta = \pi^\ast \omega$.
				\end{itemize}
			\end{column}
			\begin{column}{.4\linewidth}
				\begin{center}
					\includestandalone[width=1\textwidth]{Pictures/Frame_prequantum}
				\end{center}	
			\end{column}
		\end{columns}
	\pause \pause \vfill
	\pause
	The Lie algebra map
	\begin{displaymath}
		\big(C^\infty(M),\{\cdot,\cdot\}_\omega\big) \xrightarrow{\Psi_\omega} 
		\big(\Gamma(TM\oplus\mathbb{R}),[\cdot,\cdot]_\omega\big)
	\end{displaymath}
	%
	\vfill
	factors through the Lie algebra of \emph{\alert{infinitesimal quanto-morphisms}}
	%
	\pause
	\begin{displaymath}
		Q(P,\theta)= \left \lbrace 
			Y \in \mathfrak{X}(P) 
			~ \Big\vert ~
			\mathcal{L}_Y \theta = 0
			\right \rbrace~.
	\end{displaymath}

@@

	\hfill \onslide<5->{\hyperlink{frame:scopeTalk}{\color{gray}$[<<]$}}%\beamerskipbutton{}}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

	Consider a \emph{\bf{first-order Lagrangian Field Theory}}:
	\begin{itemize}
		\item<1-> {\color{blue} Kinematics} is encoded by a {\color{blue} \emph{configuration bundle}};
		\item<2-> {\color{red} Dynamics} is encoded by a {\color{red} \emph{Lagrangian density}};
		\item<3-> The {\color{orange} multicontangent bundle} is naturally $n$-plectic with $\omega := d \theta$ and \\
		$\theta \in \Omega^n(\Lambda)$ the tautological $n$-form
		{\small
		\begin{displaymath}
			\left[ \iota_{u_1 \wedge \ldots \wedge u_n} \theta \right]_\eta 
			= \iota_{u_1 \wedge \ldots \wedge u_n} \pi^\ast \eta 
			\qquad \qquad \forall \eta \in \wedge^n T^* E \, , \: \forall u_i \in T_\eta (\wedge^n T^* E) 		
		\end{displaymath}
		}
		\item<4-> The multisymplectic $(n+1)$-form is pulled back to $J^1 E$ through the \emph{Legendre transformation} $\mathbb{F}\mathcal{L} J^1 E \to (J^1 E)^\circledast$
    given intrinsically by
    {\small
		\begin{displaymath}
        		\Big\langle \mathbb{F}\mathcal{L}(\gamma),\gamma'\Big\rangle=
        		\mathcal{L}\Big(\gamma\Big) + \frac{ d}{ d \epsilon}\mathcal{L}\Big(\gamma +\epsilon(\gamma'-\gamma)\Big)\Bigg\vert_{\epsilon=0} \qquad \forall \gamma,\gamma' \in J^1 E
    		\end{displaymath}
    }

    %
	\end{itemize}
	%
	\vfill
	\hfill \onslide<4->{\hyperlink{frame:hystMS}{\color{gray}$[<<]$}}%\beamerskipbutton{}}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[t]{Vinogradov Algebroids}
	\begin{defblock}[Vinogradov algebroid (higher Courant)]
		\includestandalone[width=0.95\textwidth]{Pictures/Frame_vinogradov}		
	\end{defblock}
	\vfill
	\begin{itemize}
		\item<6-> $(n=1) ~ \Rightarrow$ standard twisted {\small \color{UniGreen} (by a closed $2$-form)}  \emph{Lie algebroid};
		\item<7-> $(n=2) ~ \Rightarrow$ standard twisted {\small \color{UniGreen} (by a closed $3$-form)} \emph{Courant algebroid};
	\end{itemize}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{Vinogradov $L_\infty$-algebra}\label{frame:VinoLinfty}
	Vin. alg.oids 
	$\quad\simeq\quad$ 	
	$NQ$-manifolds {\small($L_\infty$-algebroids)}
	$\quad\Rightarrow\quad$ 
	{\small Associated} $L_\infty$-algebra
	%
	\pause
	\begin{defblock}[Vinogradov $L_\infty$-algebra (Zambon) \cite{Zambon2012}]
		\includestandalone[width=0.95\textwidth]{Pictures/Frame_vinogradov-Linfty}	
	\end{defblock}
	\vfill
	\only<6->{
		\tikz[overlay,remember picture]
		{
			\node[rounded corners,
                 fill=gray!1,draw=gray!30,anchor=base]            
            	 (base) at ($(current page.south)+(0,.5)$) [rotate=-0,text width=10cm,align=center] { \footnotesize{\color{gray}{
            	 $e_i = \binom{X_i}{\alpha_i} \in \mathfrak{X}(M)\oplus \Omega^{n-1}(M)$ 
            	 \quad~,\qquad
            	 $f_i \in \bigoplus_{k=0}^{n-2}\Omega^k(M)$.
            	 }}};
		}			
	}
	\hfill \onslide<6->{\hyperlink{frame:VinoDetails}{\color{gray}$[>>]$}}%\beamerskipbutton{}}

@@

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[t,fragile,shrink=20]{Details on the Vinogradov $L_\infty$-algebra Construction}\label{frame:VinoDetails}
	%
	\begin{enumerate}[<+-| alert@+>]
		\setlength\itemsep{2em}
		\item Consider the \emph{Graded manifold}
		 $$ Q = T^\ast[n] T[1] M $$
		 call $\mathcal{C}$ the corresponding algebra of functions.
		\item $Q$ is a (shifted) contangent, hence it's naturally Poisson
		 $$ (\mathcal{C},\cdot,\{\cdot,\cdot\}) \qquad \text{is a Poisson $n$-algebra}~;.$$
		\item consider coordinates on $Q\sim (x^i,v_i, P_i, p_i)$ with
			$$ |x^i|=0,~ |v^i|=1,~ |P_i|=n, ~ |p_i|=n-1 ~;$$
		\item The degree $n+1$ function $\mathcal{S}= \sum_i v_i P_i$ encoding the de Rham differential satisfies
			$\{\mathcal{S},\mathcal{S}\}=0$. Hence
			 $$ (\mathcal{C}[n],\delta=\{\mathcal{S},\cdot\},\{\cdot,\cdot\}) \qquad \text{is a DGLA}~;$$
		\item Apply Getzler \cite{Getzler1991} and Fiorenza-Manetti \cite{Fiorenza2006}  \emph{Cone costruction} to the DGLA, one gets a $L_\infty$ algebra on the non-negative truncation of $\mathcal{C}[n]$ given by
		\begin{displaymath}
			\begin{tikzcd}[column sep= small,row sep=small]
				0 \ar[r,] & C^\infty(M) \ar["d",r] &
				\Omega^1(M) \ar["d",r] & \cdots \ar[r]
				& \Omega^{n-2}(M) \ar[r]& \Omega^{n-1}(M)\oplus \mathfrak{X}(M) \ar[r] & 0
			\end{tikzcd}
		\end{displaymath}			 
			via derived brackets.	
	\end{enumerate}
	\hfill \onslide<5->{\hyperlink{frame:VinoLinfty}{\color{gray}$[<<]$}}%\beamerskipbutton{}}
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

@@

	}
	\hfill \onslide<4->{\hyperlink{frame:thankyouslide}{\color{gray}$[\beamerskipbutton{}]$} }

@@

%##################################################################################
\begin{frame}
	\begin{center}
	\Huge\emph{Outlook}
	\end{center}
\end{frame}
\addtocounter{framenumber}{-1}
%##################################################################################

%-------------------------------------------------------------------------------------------------------------------------------------------------
\begin{frame}[fragile]{A bigger picture about Multisymplectic observables}
	OUTLOOK. La big picture. 
\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%-------------------------------------------------------------------------------------------------------------------------------------------------
	\begin{frame}{}
	\label{frame:thankyouslide}
		\vfill
	  \centering 
	  {\Huge\color{red} 
	  \emph{Thank you for your attention!}}
		\vfill
		%
		\centering
		\fbox{\includegraphics[width=.4\textwidth]{Pictures/arxiv-cover}}
	\end{frame}
%-------------------------------------------------------------------------------------------------------------------------------------------------

%------------------------------------------------------------------------------------------------
% APPENDIX
%------------------------------------------------------------------------------------------------
\appendix
\renewcommand{\insertframenumber}{\Alph{framenumber}}

%-------------------------------------------------------------------------------------------------------------------------------------------------
\section{Appendix}
%------------------------------------------------------------------------------------------------
%##################################################################################
\begin{frame}
	\begin{center}
	\Huge\emph{Supplementary Material}
	\end{center}
\end{frame}
\addtocounter{framenumber}{-1}
%##################################################################################
